---
title: "Exploratory Data Analysis"
author: "K Iwasaki"
date: "September 15, 2017"
output:
  pdf_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('C:/Users/K/Desktop/Project/JOB_HUNTING_MADE_EASY/startup_db_recommendation/explatory_data_analysis')
data = read.csv("article_after_processing10.csv")
library(plyr)
library(dplyr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)

```
# Set-up

Most of columns are categorical variables.

```{r, fig.height = 4, fig.show='hold', fig.align='center'}
drops = c("title", "link", "excerpt", "Company", "money_raised", "linkedin_link",
          "Company_at_Linkedin", "Specialties", "Website", "Location",
          "zip_code","State",
          "Description", "Also.viewed", "Industry")
data = data[, !(names(data) %in% drops) ]

summary(data[, 1:13 ])

# check columns 1:13. Columns 13: have same format.
str(data[, 1:13 ])

# show columns with na
na = lapply(data, function(x) sum(ifelse(is.na(x) | x == "" | x == "not found", TRUE, FALSE)))
na[na > 0]

```

# Univariate Analysis

- Funding round

As intended, most companies are in Series B and Series C. Need to merge "series C" and "Series C".

```{r, fig.height = 3, fig.show='hold', fig.align='center'}
# counts = table(data$funding_round)
# counts

data$funding_round[data$funding_round == "series C"] = "Series C"

# remove the level does not occur ("series C")
data$funding_round = factor(data$funding_round)

counts = table(data$funding_round)
counts

prop.table(table(data$funding_round))

barplot(counts, main = "Funding Round", las = 2)

```
- Money raised

Most of companies raised money under $100M. We observe some outliers: Magic Leap, Pivotal, GitHub, and Opendoor.com.

```{r}
summary(data$money_raised_float)

hist(data$money_raised_float, breaks = 40, main = "Money Raised")

# check the outliers
data[data$money_raised_float > 200, c("CompanyName", "funding_round", "CompanySize","money_raised_float")]

```

- Number of Employees

```{r, fig.height = 3.5, fig.show='hold', fig.align='center'}
# counts = table(data$CompanySize)
# counts

# str(data$CompanySize)

# clean up - factors
data$CompanySize = revalue(data$CompanySize, c("Nov-50"="11-50", "10-Jan"="1-10"))

# clean up - the level orders
data$CompanySize = factor(data$CompanySize, levels = c("1-10", "11-50","51-200","201-500", "501-1000",
                                           "1001-5000","5001-10,000","10,001+"))

counts = table(data$CompanySize)
counts

barplot(counts, main = "Company size", las=2)

```

- When Companies Are Founded

There are some companies founded before 2000. I suspect Hillshire Brands, founded in 1939, is a startup.

```{r}
counts = table(data$Founded)
counts

data[data$Founded < 2000, c("Founded", "CompanyName", "funding_round", 
                                      "CompanySize","money_raised_float")]

```

- Country Companies Are Based In

Since I collected startups from TechCrunch, the US based news outlet, it turns out 77% startup in the dataset are based in the US. This might also be because the US produces the largest number of startups.

```{r}
counts = table(data$Country)
counts

prop.table(counts)

```

- City


```{r, fig.height = 4, fig.show='hold', fig.align='center'}
detach(package:plyr)
data %>%
  group_by(City) %>%
  summarize(n = n()) %>%
  arrange(desc(n)) %>%
  filter(City != "") %>%
  slice(1:10) %>%
  ggplot(., aes(x = reorder(City, -n), y = n)) +
  geom_bar(stat = "identity") +
  ggtitle("10 Top Cities Startups are based in") +
  theme(plot.title = element_text(hjust = 0.5, size=14)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(axis.title.x=element_blank()) +
  ylab("Number of Companies")
  
```


# Bivariate Analysis

- Funding_round x money_raised

```{r, fig.height = 3, fig.show='hold', fig.align='center'}

ggplot(data = data, aes(x= funding_round, y = money_raised_float)) +
  geom_jitter()

```

The table below shows mean and standard deviaiton of money raised for companies in each funding round. It makes sense that the mean increases as funding round progresses. Series E has lower mean than Series D. This might be because Series E is more of extension of Series D to sustain funding and not a funding round to drive a company to next level. Also note that stand devidations are quite larger for each round.

```{r, fig.height = 6, fig.show='hold', fig.align='center'}

par(mfrow=c(2,1))

money = data %>%
  group_by(funding_round) %>%
  summarize(sum = sum(money_raised_float), mean = mean(money_raised_float), sd = sd(money_raised_float))
money

counts = money$sum
names(counts) = money$funding_round
barplot(counts, las = 2, main = "Total Money Raised by Funding Round")

counts = money$mean
names(counts) = money$funding_round
barplot(counts, las = 2, main = "Average Money Raised by Funding Round")



```

- Company size x money_raised

```{r, fig.height = 3, fig.show='hold', fig.align='center'}

ggplot(data = data, aes(x= CompanySize, y = money_raised_float)) +
  geom_jitter(alpha = 0.6) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


```{r, fig.height = 3, fig.show='hold', fig.align='center'}
money = data %>%
  group_by(CompanySize) %>%
  summarize(mean = mean(money_raised_float), sd = sd(money_raised_float))

counts = money$mean
names(counts) = money$CompanySize
barplot(counts, las = 2, main = "Average Money Raised by Company Size")

```

# Location

```{r, fig.height = 5, fig.width = 8, fig.show='hold', fig.align='center', message=F, warning=F}
# Set the center of map

all_states <- map_data("state")

ggplot() + geom_polygon(data = all_states, aes(x=long, y = lat, group = group), 
                        colour="white", fill="grey10") + 
  coord_fixed(1.3) +
  geom_jitter(data = data, mapping = aes(
    x = longitude, y = latitude, color = "Orangered" ,size = money_raised_float),
             alpha = 0.4, width = 0.7, height = 0.7) +
  scale_size(range = c(2, 12)) +
  ggtitle("Company Location with Money Raised") +
  theme(plot.title = element_text(hjust = 0.5, size=18)) +
  labs(x=NULL, y=NULL) +
  theme(panel.border = element_blank())


```





